---
title: "Reinforcing learning increasingly shapes memory specificity from childhood to adulthood"
subtitle: "Data Science Presentation"
author: "Beatriz Goncalves"
institute: "Houghton University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
# Load your data here
e1_learning_data <- read.csv("/cloud/project/data/e1_learning_data.csv")
   
e2_learning_data <- read.csv("/cloud/project/data/e2_learning_data.csv")
   
e1_memory_data <- read.csv("/cloud/project/data/e1_memory_data.csv")
   
e1_memory_data <- read.csv("/cloud/project/data/e1_memory_data.csv")
  
e1_rl_data <- read.csv("/cloud/project/data/e1_rl_data.csv")
   
e1_sub_ages <- read.csv("/cloud/project/data/e1_sub_ages.csv")
   
e2_memory_data <- read.csv("/cloud/project/data/e2_memory_data.csv")
   
e2_rl_data <- read.csv("/cloud/project/data/e2_rl_data.csv")
  
e2_sub_ages <- read.csv("/cloud/project/data/e2_sub_ages.csv")
   
```


```{r}
style_xaringan(title_slide_background_image = "https://images.unsplash.com/photo-1517673132405-a56a62b18caf?q=80&w=2076&auto=format&fit=crop&ixlib=rb-4.1.0&ixid=M3wxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8fA%3D%3D"
)
```

---
class: center, middle

## Research Question: What is the role of cognitive development in the interaction between reinforcement learning and memory specificity?


---

# Background Information
-  Kate Nussenbaum and Catherine Hartley conducted a study with New York University on how reinforcement learning increasingly shapes memory specificity from childhood to adulthood. The data was collected through two multi-session online experiments. 
- In the first session of each experiment, participants completed a reinforcement-learning task in which either specific examples or broader stimulus categories determined the rewards associated with different actions. 
- In the second session, which was administered one week later, participants completed a memory test in which we probed the specificity of their memories for the stimuli encountered during learning.


---
class:center, middle

# Learning Data Comparison Table

---

```{r, echo = FALSE}

e1_learning_data <- e1_learning_data %>% mutate(experiment = "Experiment 1")
e2_learning_data <- e2_learning_data %>% mutate(experiment = "Experiment 2")

learning_sum <- bind_rows(e1_learning_data, e2_learning_data)

kable(head(learning_sum), format = "html")



```
---

-The table demonstrates how experiment 1 and experiment 2 are mainly similar, but the main difference is that in experiment 1 they use a reward approach that is associated with broader stimuli. The reasoning behind this reward approach is to enhance learning and memory in those categories so the participants are able to remember the information for experiment 2. 


---

# Is there any relation between one's age and one's memory in experiment 2?

---

```{r, echo=FALSE}

library(ggplot2)
library(dplyr)


memory_age2 <- merge(e2_memory_data, e2_sub_ages, by = "subject_id")

memory_age2_clean <- memory_age2 %>%
  select(age, mem_acc) %>%
  filter(!is.na(age), !is.na(mem_acc), is.finite(age), is.finite(mem_acc))


ggplot(memory_age2_clean, aes(x = age, y = mem_acc)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "blue") +
  labs(title = "Memory Accuracy vs Age (Experiment 2)",
       x = "Age",
       y = "Memory Accuracy") +
  theme_minimal()


```


---




- According to the visualization, you notice a slightly negative correlation between one's memory and one's age in experiment 2. Meaning, that according to the plot, older participants tend to have slightly lower memory accuracy. These results may come to a shock because overall some of the older participants perform just as well as the younger ones.

---



# How is reinforcement learning influential in experiment 1 and experiment 2?

---


```{r, echo=FALSE}

e1_rl_data$experiment <- "Experiment 1"
e2_rl_data$experiment <- "Experiment 2"

e1_rl_data <- e1_rl_data %>%
  select(subject_id, block_order, trial, points, experiment)
e2_rl_data <- e2_rl_data %>%
  select(subject_id, block_order, block_condition, trial, points)
rl_data <- bind_rows(e1_rl_data, e2_rl_data)

summary <- rl_data %>%
  group_by(experiment, points) %>%
  summarise(avg_points = mean(points, na.rm = TRUE), .groups = "drop")

ggplot(summary, aes(x = experiment, y = avg_points, fill = experiment)) + 
  geom_boxplot() +
  labs(title = "Average Points per Subject by Experiment",
       x = "Experiment",
       y = "Average Points") +
  theme_minimal()
```


---



```{r, echo=FALSE}
rl_data %>%
  group_by(experiment, trial) %>%
  summarise(mean_points = mean(points, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = trial, y = mean_points, color = experiment)) +
  geom_line(linewidth =1) +
  labs(title = "Points /over Trials Experiment", 
       x = "Trial",
       y = "Mean Points") +
  theme_minimal()
```

---

- The differences between experiment 1 and experiment 2 is that experiment 2 shows higher average points, which suggests improved task design. Experiment 2 has developmental differences, which can be useful to the older participants. In Experiment 1, the scores are similar, but in this experiment it prioritized applying reinforcement cues effectively.



---

# What can one conclude about the effectiveness of reinforcement learning when shaping memory across all ages?
---
- This tells us that reinforcement learning increasingly shapes memory specificity from childhood to adulthood by suggesting that as cognitive systems mature, people become better at selectively encoding and applying information that matters for rewards.
---

# The End

---